---
title: "Data analysis"
author: "Lainy VonBank"
date: "09.12.2022"
output: html_document
---
# Load packages and data
```{r, message=FALSE, warning=FALSE}
if(!require(tidyverse)){install.packages("tidyverse")}
if(!require(dplyr)){install.packages("dplyr")}
if(!require(ggpubr)){install.packages("ggpubr")}
if(!require(reshape)){install.packages("reshape")}
if(!require(rstatix)){install.packages("rstatix")}
if(!require(ggrepel)){install.packages("ggrepel")}
if(!require(ggsci)){install.packages("ggsci")}
if(!require(scales)){install.packages("scales")}
if(!require(heatmaply)){install.packages("heatmaply")}
if(!require(Cairo)){install.packages("Cairo")}
if(!require(factoextra)){install.packages("factoextra")}
if(!require(RColorBrewer)){install.packages("RColorBrewer")}
if(!require(tidyverse)){install.packages("tidyverse")}
if(!require(reshape)){install.packages("reshape")}
if(!require(Cairo)){install.packages("Cairo")}
if(!require(ggpubr)){install.packages("ggpubr")}
if(!require(rstatix)){install.packages("rstatix")}
if(!require(ggforce)){install.packages("ggforce")}
library(forcats)
```


# Volcano Plot
```{r}
list.files("cleanup/")
dfs <- read_csv("cleanup/combined_pmol_mg_tissue.csv")
dfs <- dfs %>%
  unite("Condition",2:3,remove=T)
dfs$Condition <- gsub("HNF4a ", "", dfs$Condition)
View(dfs)
dfs <- dfs[-3] #remove weight column

cold <- dfs %>%
  filter(str_detect(Condition,"cold"))
cold$Condition <- gsub("_cold","",cold$Condition)
RT <- dfs %>%
  filter(str_detect(Condition,"RT"))
RT$Condition <- gsub("_RT","",RT$Condition)

Z <- cold
X = matrix(NA, ncol = 3, nrow = ncol(Z))
colnames(X) <- c("FC", "p", "sig")
rownames(X) <- colnames(Z)

p_value = as.numeric()
  
  for (j in 3:ncol(Z)) {                               
    X[j,1] = round(colMeans(Z[Z$Condition == "KO",j],na.rm = T)/colMeans(Z[Z$Condition == "Control",j], na.rm = T), 4)
    
      if(table(is.na(Z[Z$Condition == "KO",j]))[1] > 1 && table(is.na(Z[Z$Condition == "Control",j]))[1] > 1){
        X[j,2] = round(-log10(t.test(Z[Z$Condition == "KO",j], Z[Z$Condition == "Control",j], na.rm = T)$p.val), 4)
      }else(X[j,2] == 0.9)
      
      if(table(is.na(Z[Z$Condition == "KO",j]))[1] > 1 && table(is.na(Z[Z$Condition == "Control",j]))[1] > 1){
        p_value[j] = t.test(Z[Z$Condition == "KO",j], Z[Z$Condition == "Control",j], na.rm = T)$p.val
      }else(X[j,2] == 0.9)
  }

  op <- p.adjust(p_value[3:length(p_value)], method = 'fdr')
  Y1 <- data.frame(X[3:nrow(X),])
  Y1$sig <- op < 0.15
  
  Y1 <- filter(Y1, FC != 0)
  Y1 <- filter(Y1, FC != Inf)
  View(Y1) 
  
  Y1$FC <- log2(Y1$FC)
  
  Y1$diffexpressed <- "NO"
  Y1$diffexpressed[Y1$FC > 0.0 & Y1$sig == T] <- "UP"
  Y1$diffexpressed[Y1$FC < 0.0 & Y1$sig == T] <- "DOWN"
  
  Y1$diffexpressed = factor(Y1$diffexpressed, levels = c("NO", "UP", "DOWN"))
  
  Y1$delabel <- NA
  Y1$delabel[Y1$diffexpressed != "NO"] <- rownames(Y1)[Y1$diffexpressed != "NO"]
  
  toMatch <- c("20:4","20:5","22:5", "22:6")
  Y1$delabel <- ifelse(grepl(paste(toMatch,collapse="|"), 
                             Y1$delabel),Y1$delabel,NA)
  
  Y1$label <- NA
  Y1$label[Y1$sig == T] <- "sig"
  Y1$label[Y1$sig == F] <- "not"
  
  n_sig <- table(Y1$sig)
  n_direction <- table(Y1$diffexpressed)
 
  p0 <-ggplot(data=Y1, aes(x=FC, y=p, col=diffexpressed)) +
        geom_point() + 
        scale_x_continuous(limits = c(max(Y1$FC)*-3, max(Y1$FC)*3)) +
        scale_y_continuous(limits = c(0, max(Y1$p*1.1))) +
        theme_minimal() +
        #geom_label_repel(aes(x=FC, y=p, label=delabel), min.segment.length = 0, size = 2.5, family = "Arial", max.overlaps = Inf) +
        scale_color_manual(values=c("gray", "#E64B35FF", "#4DBBD5FF"), name = "Liver Lipids HNF4a KO vs Control Cold",
                           breaks = c("NO", "UP", "DOWN"), 
                           labels = c("FDR > 0.15", "Increased", "Decreased")) +
        ggtitle(names(Z)) +
        ylab("-log10(p-value)") +
        xlab("log2(Fold change [KO/Control])") +
        theme(legend.position = "right", text = element_text(size = 12, family = "Arial"))
  
  CairoPDF(file = paste0("graphs/volcano_cold_FDR_0.15.pdf"), height = 8, width = 12,)
  print(p0)
  #dev.off() 
  
  colnames(Y1)[1:3] <- c("log2(FC)", "log10(p)", "sig_FDR")
  Y1$FC = round(2^Y1[,1], 4)
  Y1$p_val = round(10^Y1[,2], 4)
  
  write_csv(Y1, "volcano_cold_fdr0.15.csv")
rm(list=ls())
```


# Lipid class distribution
```{r}
list.files("cleanup/")
mydata <- read_csv("cleanup/combined_pmol_mg_tissue.csv")
mydata[mydata == 0] <- NA
mydata<- mydata %>%
  unite("Condition",2:3,remove=T)
mydata$Condition <- gsub("HNF4a ", "", mydata$Condition)
mydata1 <- mydata[-c(3)] #remove tissue weight


mydata1_PL <- mydata1[,c(1:2,grep("PC", colnames(mydata1)),grep("PE", colnames(mydata1)),grep("PI", colnames(mydata1)),grep("PG", colnames(mydata1)),grep("PS", colnames(mydata1)), grep("PMe", colnames(mydata1)),grep("BMP", colnames(mydata1)), grep("DG", colnames(mydata1)), grep("PA", colnames(mydata1)))]
colnames(mydata1_PL) <- gsub("/", "_", colnames(mydata1_PL))

mydata1_Cer_other <- mydata1[,c(1:2,grep("Cer", colnames(mydata1)),grep("SM", colnames(mydata1)), grep("FAH", colnames(mydata1)))]

mdata1_FA <- mydata1[,c(1:2, grep("FA", colnames(mydata1)))]
mdata1_FA <- mdata1_FA[-grep("FAHFA", colnames(mdata1_FA))]


mydata1_TG <- mydata1[,c(1:2,grep("TG", colnames(mydata1)))]

```

#Phospholipids
```{r}
mdat1 <- mydata1_PL 
mdat1[is.na(mdat1)] <- 0

mdat2 <- reshape2::melt(mdat1, id = c(1:2))
mdat2$variable <- gsub("\\.1", "", mdat2$variable)
view(mdat2)
mdat2$variable <- gsub(".*PC ", "", mdat2$variable)
mdat2$variable <- gsub(".*PE ", "", mdat2$variable)
mdat2$variable <- gsub(".*PI ", "", mdat2$variable)
mdat2$variable <- gsub(".*PS ", "", mdat2$variable)
mdat2$variable <- gsub(".*PG ", "", mdat2$variable)
mdat2$variable <- gsub(".*PA ", "", mdat2$variable)
mdat2$variable <- gsub(".*BMP ", "", mdat2$variable)
mdat2$variable <- gsub(".*PMeOH ", "", mdat2$variable)
mdat2$variable <- gsub(".*DG ", "", mdat2$variable)

mdat2_FA <- str_split_fixed(mdat2$variable, "_", 2)
mdat3 <- cbind(mdat2, mdat2_FA)

mdat3_FA1 <- mdat3[,1:5]
mdat3_FA2 <- mdat3[,c(1:4,6)]

names(mdat3_FA1)[5] <- "FA"
names(mdat3_FA2)[5] <- "FA"

mdat4 <- bind_rows(mdat3_FA1, mdat3_FA2)
#check that the number of rows equals the sum of the the mdat3FA1+2
View(mdat4)
mdat4$FA <- gsub("e", " ", mdat4$FA)

chains <- c("12:0", "14:0","15:0", "16:0", "16:1","17:0","17:1", "18:0", "18:1", 
            "18:2", "18:3","19:0", "20:0" ,"20:3", "20:4", "20:5", "22:4", 
            "22:5", "22:6", "24:1")
mdat4_2 <- mdat4[mdat4$FA %in% chains,]
#Save a csv for each lipid class
write_csv(mdat4_2, "acyl_chain/acyl_PL.csv")
rm(list = ls())
```

#Ceramides/other
```{r}
mdat1 <- mydata1_Cer_other
mdat1[is.na(mdat1)] <- 0

mdat2 <- reshape2::melt(mdat1, id = c(1:2))
mdat2$variable <- gsub("\\.1", "", mdat2$variable)
view(mdat2)
mdat2$variable <- gsub(".* d", "", mdat2$variable)
mdat2$variable <- gsub(".* t", "", mdat2$variable)
mdat2$variable <- gsub("SM ", "", mdat2$variable)
mdat2$variable <- gsub(".*FAHFA ", "", mdat2$variable)

mdat2_FA <- str_split_fixed(mdat2$variable, "_", 2)
mdat3 <- cbind(mdat2, mdat2_FA)

mdat3_FA1 <- mdat3[,1:5]
mdat3_FA2 <- mdat3[,c(1:4,6)]

names(mdat3_FA1)[5] <- "FA"
names(mdat3_FA2)[5] <- "FA"

mdat4 <- bind_rows(mdat3_FA1, mdat3_FA2)
#check that the number of rows equals the sum of the the mdat3FA1+2
View(mdat4)

chains <- c("12:0", "14:0","15:0", "16:0", "16:1","17:0","17:1", "18:0", "18:1", 
            "18:2", "18:3","19:0", "20:0" ,"20:3", "20:4", "20:5", "22:4", 
            "22:5", "22:6", "24:1")
mdat4_2 <- mdat4[mdat4$FA %in% chains,]
#Save a csv for each lipid class
write_csv(mdat4_2, "acyl_chain/acyl_Cer.csv")
rm(list = ls())
```

#FA/AC
```{r}
mdat1 <- mdata1_FA
mdat1[is.na(mdat1)] <- 0

mdat2 <- reshape2::melt(mdat1, id = c(1:2))
mdat2$variable <- gsub("\\.1", "", mdat2$variable)
view(mdat2)
mdat2$variable <- gsub("FA ", " ", mdat2$variable)
mdat2$FA <- mdat2$variable

chains <- c("12:0", "14:0","15:0", "16:0", "16:1","17:0","17:1", "18:0", "18:1", 
            "18:2", "18:3","19:0", "20:0" ,"20:3", "20:4", "20:5", "22:4", 
            "22:5", "22:6", "24:1")
mdat4_2 <- mdat2[mdat2$FA %in% chains,]
#Save a csv for each lipid class
write_csv(mdat4_2, "acyl_chain/acyl_FA.csv")
rm(list = ls())
```


#Triglycerides (separated)
```{r}
#For separated acyl chains
mdat1 <- mydata1_TG
mdat1[is.na(mdat1)] <- 0

mdat2 <- reshape2::melt(mdat1, id = c(1:2))
mdat2$variable <- gsub("\\.1", "", mdat2$variable)

mdat2$variable <- gsub("TG ", "", mdat2$variable)

mdat2_FA <- str_split_fixed(mdat2$variable, "_", 3)
mdat3 <- cbind(mdat2, mdat2_FA)

mdat3_FA1 <- mdat3[,1:5]
mdat3_FA2 <- mdat3[,c(1:4,6)]
mdat3_FA3 <- mdat3[,c(1:4,7)]
names(mdat3_FA1)[5] <- "FA"
names(mdat3_FA2)[5] <- "FA"
names(mdat3_FA3)[5] <- "FA"
mdat4 <- bind_rows(mdat3_FA1, mdat3_FA2, mdat3_FA3)


View(mdat4)
chains <- c("12:0", "14:0","15:0", "16:0", "16:1","17:0","17:1", "18:0", "18:1", 
            "18:2", "18:3","19:0", "20:0" ,"20:3", "20:4", "20:5", "22:4", 
            "22:5", "22:6", "24:1")
mdat4_2 <- mdat4[mdat4$FA %in% chains,]
#Save a csv for each lipid class
write_csv(mdat4_2, "acyl_chain/Jain22/acyl_TG.csv")
rm(list = ls())
```

#Segmented Bar graph acyl chain sums
```{r}
list.files()
N1 <- read_csv("acyl_chain/acyl_Cer.csv")
N2 <- read_csv("acyl_chain/acyl_PL.csv")
N3 <- read_csv("acyl_chain/acyl_FA.csv") 
N4 <- read_csv("acyl_chain/acyl_TG.csv")

mdat_combo <- rbind(N1,N2,N3,N4)


mdat5 <- aggregate(mdat_combo$value, by=list(Name=mdat_combo$Name, Temp=mdat_combo$Condition, FA=mdat_combo$FA), FUN=sum) #sum acyl chain per sample
View(mdat5)


mdat6 <- str_split_fixed(mdat5$`FA`,":", 2)
mdat_class <- cbind(mdat5,mdat6)
colnames(mdat_class)[colnames(mdat_class) == 'x'] <- 'value'
View(mdat_class)
mdat_class$`2` <- as.numeric(mdat_class$`2`)

mdat_class$class[mdat_class$`2` == 0] <- "SFA"
mdat_class$class[mdat_class$`2` == 1] <- "MUFA"
mdat_class$class[mdat_class$`2` > 1] <- "PUFA"

Zrt <- mdat_class[str_detect(mdat5$Temp,"RT"),]
Zrt$Temp <- gsub("_RT","",Zrt$Temp)
Zrt$class = factor(Zrt$class, levels = c("SFA", "MUFA", "PUFA"),ordered=T)

Zc <- mdat_class[str_detect(mdat5$Temp,"cold"),]
Zc$Temp <- gsub("_cold","",Zc$Temp)
Zc$class = factor(Zc$class, levels = c("SFA", "MUFA", "PUFA"),ordered=T)

p0 <- ggplot(Zrt, aes(x=Temp, y=value, fill=class)) + 
    geom_bar(position="fill", stat="identity")

p1 <- ggpar(p0, legend = "right", legend.title = "Saturation Class", 
                xlab = "RT", ylab = "% Abundance")
  p2 <- p1 + theme_bw() 

CairoPDF(file = paste0("acyl_chain/segmented_RT.pdf"), height = 8, width = 12,)
  print(p2)


```



#Triglycerides (composite). This incorporates TGs that did not have acyl chains fully annotated, but can still gather information on the number of double bonds. 
```{r}
mdat1 <- mydata1_TG 
mdat1[is.na(mdat1)] <- 0

mdat2 <- reshape2::melt(mdat1, id = c(1:2))
mdat2$variable <- gsub("\\.1", "", mdat2$variable)

mdat2$variable <- gsub(".*TG ", "", mdat2$variable)

mdat2_FA <- str_split_fixed(mdat2$variable, "_", 3)
mdat3 <- cbind(mdat2, mdat2_FA)

#For composite acyl chains 
mdat3_DB1 <- str_split_fixed(mdat3$`1`,":", 2)
mdat3_DB2 <- str_split_fixed(mdat3$`2`,":", 2)
mdat3_DB3 <- str_split_fixed(mdat3$`3`,":", 2)
mdat3_DB <- cbind(mdat3, mdat3_DB1, mdat3_DB2, mdat3_DB3)
mdat3_DB[,8:ncol(mdat3_DB)] <- sapply(mdat3_DB[,8:ncol(mdat3_DB)], as.numeric)
mdat3_DB$sumacyl <- rowSums(mdat3_DB[,c(8,10,12)], na.rm=T)
mdat3_DB$sumDB <- rowSums(mdat3_DB[,c(9,11,13)], na.rm=T)
mdat3_DB$sumDB <- as.character(mdat3_DB$sumDB)
mdat3_DB <- mdat3_DB %>% unite("sumCOMP",14:15, sep = ":", remove = F)

mdat5 <- aggregate(mdat3_DB$value, by=list(Temp=mdat3_DB$Condition, DB=mdat3_DB$sumDB), FUN=sum)
colnames(mdat5)[colnames(mdat5) == 'x'] <- 'value'
View(mdat5)

mdat5$DB <- as.numeric(mdat5$DB)
mdat5$class[mdat5$DB < 2] <- "0-1"
mdat5$class[mdat5$DB == 2] <- "2"
mdat5$class[mdat5$DB == 3] <- "3"
mdat5$class[mdat5$DB == 4] <- "4"
mdat5$class[mdat5$DB > 4] <- "5+"


mdat5$DB <- as.numeric(mdat5$DB)
mdat5$class = factor(mdat5$class, ordered = T)
mdat5$Temp = factor(mdat5$Temp, levels = c("Control_RT","KO_RT","Control_cold", "KO_cold"), ordered = T)

p0 <- ggplot(mdat5, aes(x=Temp, y=value, fill=class)) + 
    geom_bar(position="fill", stat="identity")

p1 <- ggpar(p0, legend = "right", legend.title = "Number of double bonds", 
               ylab = "% Abundance", title = "Triglyceride Composition")
  p2 <- p1 + theme_bw() 

CairoPDF(file = paste0("acyl_chain/all_TGs_segmented_grouped.pdf"), height = 8, width = 12,)
  print(p2)

######
KOc <- mdat5[mdat5$Temp == "KO_cold",]
KOrt <- mdat5[mdat5$Temp == "KO_RT",]

pie(KOc$x, labels = KOc$DB, main = "Cold KO",
    border="white")
pie(KOrt$x, labels = KOrt$DB, main = "RT KO",
    border="white")

```


#Bar graph acyl summation RT vs cold (data from Jain et al 2022)
```{r}
list.files()
N1 <- read_csv("acyl_chain/Jain22/acyl_Cer.csv")
N2 <- read_csv("acyl_chain/Jain22/acyl_PL.csv")
N3 <- read_csv("acyl_chain/Jain22/acyl_FA.csv") 
N4 <- read_csv("acyl_chain/Jain22/acyl_TG.csv")

mdat_combo <- rbind(N1,N2,N3,N4)

mdat5 <- aggregate(mdat_combo$value, by=list(Name=mdat_combo$Name, Temp=mdat_combo$Condition, FA=mdat_combo$FA), FUN=sum) #sum acyl chain per sample
View(mdat5)

mdat5$FA = factor(mdat5$FA, ordered = T)
mdat5$Temp = factor(mdat5$Temp,levels = c("RT", "Cold"), ordered = T)

a = "FA"
b = "x"
c = "Temp"
p0 <- ggbarplot(mdat5, x = a, y = b,
            add = "mean_se", color = "black", fill = c, 
            add.params = list(width = 0.35, binwidth = 0.07),
            alpha = 0.5, position = position_dodge(0.8), size = .5) 
p1 <- ggpar(p0, palette=c("white","black"), legend = "right", legend.title = c, xlab = F, ylab = "Acyl chain sum (nmol/g tissue)") +
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1, size = 12)) 

stat.test <- mdat5 %>% 
                  group_by(FA) %>% 
                  t_test(x ~ Temp) %>% 
                  adjust_pvalue(method = "none") %>%
                  add_significance("p") %>% 
                  add_xy_position(fun="mean_se",x=a,dodge=0.8)
  stat.test$p.adj.signif <- stat.test$p.signif 
  
  p3 <- p1 + stat_pvalue_manual(
      stat.test, label = "p.adj.signif", hide.ns = T)
 
  CairoPDF(file = paste0("acyl_chain/Jain22/bar.pdf"), height = 8, width = 12,)
  print(p3)
   

```









