---
title: "HNF4a_Cleanup"
author: "Lainy VonBank"
date: "7/28/2022"
output: html_document
---

---

# Load packages
```{r}
if(!require(tidyverse)){install.packages("tidyverse")}
if(!require(dplyr)){install.packages("dplyr")}
if(!require(Cairo)){install.packages("Cairo")}
getwd()
```

# Load data output from Profinder Analysis (exported detailed csv)
```{r, message = F}
list.files()
mydata <- read_csv("PFA_feb21_ether_neg_area.csv")
Y <- mydata %>% select(grep("Name", names(mydata)), grep("Area", names(mydata)))

```


# Blank subtraction 
Subtracting the processed blank height values from samples is optional as it would apply to all samples.
The primary reason to perform the subtraction is if there are suspected lipid contaminants that are the main/sole reason for presence of certain peaks. In such a case, the blank subtraction may help elucidate this by taking height values to ~0. 

To quickly screen for such such a possibility, run the following code. 
```{r}
# Specify the name of the process blank column below after the subtraction
a <- as.name(colnames(select(Y,matches("PB"))))
print(a)
# Copy this into the second part of the equation below. Make sure it is the PB

Y1 <- mutate(Y, sub_height = `Height (avg)` - `[Area] liver_PB_neg_02172021_116`)
# Set NA values to 0
Y1$sub_height <- replace_na(Y1$sub_height, 0)
print(Y1[Y1$sub_height < 0,]$Name)
# If any values are less than 0 in Y1$sub_height, that lipid(s) is noise or a contaminant and excluded
Y1 <- Y1[!Y1$sub_height < 0,]
```

Additionally, it is possible that for some variables, many, but not all, of the samples have height below the PB.
However, the 'Avg (height)' is still above the PB height. In this case, run the following code to:

-Tabulate the number of instances where PB height is greater than a sample for each variable
-Add to that the number of times PB is a value but samples have NA values
```{r}
Y1['PB_comp'] <- 0
# Note that the column values in the below code must be adjusted for different sample sets. 
# In the second for loop, the numbers should span all sample columns
# Within the if statements, the column specification should equal the column number of PB
for (j in 1:nrow(Y1)){
   for (i in 3:20) {
    if(is.na(Y1[j, 21]) == F && is.na(Y1[j, i]) == F && Y1[j, 21] >= Y1[j,i]){
      Y1[j,]$PB_comp = Y1[j,]$PB_comp + 1
    }else if(is.na(Y1[j, 21]) == F && is.na(Y1[j, i]) == T){
      Y1[j,]$PB_comp = Y1[j,]$PB_comp + 1
    }else(Y1[j,]$PB_comp = Y1[j,]$PB_comp)
  }
}
# Remove any variables where half of the samples have values of NA or less than PB
print(Y1[Y1$PB_comp >= 8,]$Name)
Y1 <- Y1[Y1$PB_comp < 8,]
```

# Remove extraneous columns
```{r}
Y2 <- Y1 %>% select(grep("Name", names(Y1)), grep("pos", names(Y1)),
                          grep("neg", names(Y1)))
# Removes PB
Y3 <- Y2 %>% select(-grep("PB", names(Y2)))
Y3 <- Y3 %>% select(-grep("2uL", names(Y3))) #removing the samples that were run at a different volume 
```

# Rename sample names
Tailor this section to filename. Note that .* means either 'from beginning' or 'to the end'.
```{r}
colnames(Y3)
colnames(Y3) <- gsub(".*liver_", "", colnames(Y3))
colnames(Y3) <- gsub("_neg.*" , "", colnames(Y3))
print(colnames(Y3))
```

# Filtering of variables with missing data
Remove lipid variables with less than a certain amount of data availability. This is deprecated. 
```{r}
# Set the minimum number of samples that must have a value for a given lipid
a = 3
missing.value = as.character()
for (i in 1:nrow(Y3)){
  if(table(is.na(Y3[i,]))[1] > a)
    missing.value[i] = "NO"
  else(missing.value[i]="YES")
}
Y3_rm = data.frame(cbind(rownames(Y3), missing.value)) 
d <- as.numeric(Y3_rm[Y3_rm$missing.value=="YES", 1])
print(Y3[d,]$Name)
d_name <- Y3[d,]$Name
Y3 <- Y3[!Y3$Name %in% d_name,]
```

# Transpose data
```{r}
Y3 <- sapply(Y3, as.character)
Y4 <- t(Y3)
Y5 <- as_tibble(Y4, rownames = "Name")
colnames(Y5) <- Y5[1,]
Y6 <- Y5[-1,]
Y6[,2:ncol(Y6)] <- sapply(Y6[,2:ncol(Y6)], as.numeric) 
```

# Lipid correlation for duplicate screening [do not use if high number of NA values]
Correlate lipids. If R=1, same peak used for both lipids. However, if high number of NAs, this will not work.  
Alternatively, run a strict exclusion below, then correlate to identify lipids to exclude. 
```{r}
# Set the minimum number of samples that must have a value for a given lipid [correlation]
a = 17
missing.value = as.character()
for (i in 1:ncol(Y6)){
  if(table(is.na(Y6[,i]))[1] > a)
    missing.value[i] = "NO"
  else(missing.value[i]="YES")
}
Y6_rm = data.frame(cbind(colnames(Y6), missing.value)) 
d <- Y6_rm[Y6_rm$missing.value=="YES", 1]
Z <- Y6[,!colnames(Y6) %in% d]   
#Z1 <- Z[-1,] 
# Now run correlation
Z2 <- Z[,2:ncol(Z)]
z = cor(Z2, use = "pairwise.complete.obs")
z[lower.tri(z, diag=TRUE)] = NA
z <- z[,2:ncol(z)]
z = as.data.frame(as.table(z))
z = z[order(-abs(z$Freq)),]
y <- z[!is.na(z$Freq)==T,]
lipid.dups = y[y$Freq>0.99999,]
lipidup.list = as.vector(lipid.dups$Var1) 
lipidup.list_2 = as.vector(lipid.dups$Var2)
# list of likely duplicate lipids (may not be all)
print(lipidup.list)  
print(lipidup.list_2)                           
```

Determine if list is acceptable to remove duplicates. 
```{r}
# Adjust which lipids to remove from 'lipidup.list'.  0 = no lipids removed. 
# For example, if the second and fourth lipids from the list need to be removed but not others:
# In brackets after lipidup.list, type '[c(2,4)]'
#rm_lipid <- lipidup.list[c(1:length(lipidup.list))]
#remove the duplicate annotation with less information on the acyl chain (eg remove TG58:7, keep TG18:0_18:1_22:6)
rm_lipid <- lipidup.list[c()] 
rm_lipid_2 <- lipidup.list_2[c(1)]
Y7 <- Y6[,!colnames(Y6) %in% rm_lipid]   
Y8 <- Y7[,!colnames(Y7) %in% rm_lipid_2]   
```

# Add sample information and produce initial output
```{r}
list.files()
sample_info <- read_csv("HNF4a_sample_prep.csv")
# Ensure that the sample ID containing column is renamed to 'Name' to match Y7
colnames(sample_info)[1] <- "Name"
# Check for any other sample id discrepancies
Y8$Name <- as.numeric(Y8$Name)
lipid_data <- merge(sample_info, Y8, by = "Name")

```
Verify data looks good and output initial merge file for further QC and analysis
```{r}
write_csv(lipid_data, "cleanup/ether_neg_area_initial.csv")
rm(list = ls())
```
Repeat for other ionization mode before continuing.

# Plot standards
```{r, message=F}
list.files("cleanup/")
pos_dat <- read_csv("cleanup/pos_height_initial.csv")
pos_dat$Name <- as.character(pos_dat$Name)
neg_dat <- read_csv("cleanup/ether_neg_height_initial.csv")
neg_dat$Name <- as.character(neg_dat$Name)
```

```{r}
pos_IS <- colnames(select(pos_dat , matches("(d7)")))
pos_IS <- append(pos_IS, colnames(select(pos_dat , matches("(d9)"))))
pos_IS <- sort(pos_IS)
neg_IS <- colnames(select(neg_dat , matches("(d7)")))
neg_IS <- append(neg_IS, colnames(select(neg_dat , matches("(d9)"))))
neg_IS <- sort(neg_IS)
plot_list <- list()

for(i in 1:length(pos_IS)){
  var_name <- paste0("`", as.name(pos_IS[i]), "`")
  p <- ggplot(data = pos_dat, aes_string(x= "Name", y = sprintf("log10(%s)", var_name))) +
             geom_boxplot() + scale_y_continuous(limit = c(3, 7)) + ggtitle(pos_IS[i])
  plot_list[[i]] <- p
}
CairoPDF("cleanup/pos_IS.pdf")
  for (i in 1:length(pos_IS)) {
    print(plot_list[[i]])
  }
dev.off()
for(i in 1:length(neg_IS)){
  var_name <- paste0("`", as.name(neg_IS[i]), "`")
  p <- ggplot(data = neg_dat, aes_string(x= "Name", y = sprintf("log10(%s)", var_name))) +
             geom_boxplot() + scale_y_continuous(limit = c(3, 7)) + ggtitle(neg_IS[i])
  plot_list[[i]] <- p
}
CairoPDF("cleanup/neg_IS.pdf")
  for (i in 1:length(neg_IS)) {
    print(plot_list[[i]])
  }
dev.off()
```

# IS normalization (semi-quantitative) ############################ This quantification is relative to the amount of internal standard injected for a representative lipid class. 


Set the dilution factor (DF) and volume multiplier here to make negative and positive mode indicative of the same injected sample amount. Volume multiplier is 5 uL injected in negative mode of 2X diluted, and 3 uL injected in positive mode if 20X diluted. 

Concentrations from Avanti SPLASH mix Lot #33707-013

```{r}
# 10 uL IS added to each sample
# Resuspended in 150 uL after evaporation
# DF = 20
# volume = 3
# Example PC = 150.6 ug/mL = 0.199  umol/mL

#Values in pmol
PC_p <- pos_dat$`PC 15:0_18:1(d7)`
PC_p_pmol <- ((150.6/753.11)*0.01*(3/150)*1000*1000*(1/20))
LPC_p <- pos_dat$`LPC 18:1/0:0(d7)`
LPC_p_pmol <- ((23.8/538.72)*0.01*(3/150)*1000*1000*(1/20))
PE_p <- pos_dat$`PE 15:0_18:1(d7)`
PE_p_pmol <- ((5.3/711.03)*0.01*(3/150)*1000*1000*(1/20))
TG_p <- pos_dat$`TG 15:0_18:1(d7)_15:0`
TG_p_pmol <- ((52.8/812.37)*0.01*(3/150)*1000*1000*(1/20))
SM_p <- pos_dat$`SM d18:1_18:1(d9)`
SM_p_pmol <- ((29.6/738.12)*0.01*(3/150)*1000*1000*(1/20))


```

Miscellaneous pos first before the IS is normalized by itself. 
```{r}
pos_dat[,grepl("^BMP" , names(pos_dat))] <- 
  (pos_dat[,grepl("^BMP" , names(pos_dat))]/PC_p)*PC_p_pmol
pos_dat[,grepl("^Cer" , names(pos_dat))] <- 
  (pos_dat[,grepl("^Cer" , names(pos_dat))]/SM_p)*SM_p_pmol
pos_dat[,grepl("^CL" , names(pos_dat))] <- 
  (pos_dat[,grepl("^CL" , names(pos_dat))]/PC_p)*PC_p_pmol
pos_dat[,grepl("^HexCer" , names(pos_dat))] <- 
  (pos_dat[,grepl("^HexCer" , names(pos_dat))]/SM_p)*SM_p_pmol
pos_dat[,grepl("^PS" , names(pos_dat))] <- 
  (pos_dat[,grepl("^PS" , names(pos_dat))]/PC_p)*PC_p_pmol
```

```{r}
pos_dat[,grepl("^ACar" , names(pos_dat))] <- 
  (pos_dat[,grepl("^ACar" , names(pos_dat))]/LPC_p)*LPC_p_pmol
#pos_dat[,grepl("^DG" , names(pos_dat))] <- 
  #(pos_dat[,grepl("^DG" , names(pos_dat))]/DG_p)
pos_dat[,grepl("^LPC" , names(pos_dat))] <- 
  (pos_dat[,grepl("^LPC" , names(pos_dat))]/LPC_p)*LPC_p_pmol
pos_dat[,grepl("^LPE" , names(pos_dat))] <- 
  (pos_dat[,grepl("^LPE" , names(pos_dat))]/LPC_p)*LPC_p_pmol
pos_dat[,grepl("^PC" , names(pos_dat))] <- 
  (pos_dat[,grepl("^PC" , names(pos_dat))]/PC_p)*PC_p_pmol
pos_dat[,grepl("^PE" , names(pos_dat))] <- 
  (pos_dat[,grepl("^PE" , names(pos_dat))]/PE_p)*PE_p_pmol
#pos_dat[,grepl("^PG" , names(pos_dat))] <- 
  #(pos_dat[,grepl("^PG" , names(pos_dat))]/PC_p)
pos_dat[,grepl("^SM" , names(pos_dat))] <- 
  (pos_dat[,grepl("^SM" , names(pos_dat))]/SM_p)*SM_p_pmol
pos_dat[,grepl("^TG" , names(pos_dat))] <- 
  (pos_dat[,grepl("^TG" , names(pos_dat))]/TG_p)*TG_p_pmol
```

#Remove lipids that did not have a standard to normalize to
```{r}
pos_dat <- pos_dat[,!grepl(c("DG"),colnames(pos_dat))]
pos_dat <- pos_dat[,!grepl(c("PG"),colnames(pos_dat))]
```


#Negative mode. 
```{r}
# 10 uL IS added to each sample
# Resuspended in 150 uL after evaporation
# DF = 2
# volume = 5

print(neg_IS)
#Values in pmol
LPE_n <- neg_dat$`LPE 18:1(d7)`
LPE_n_pmol <- ((4.9/486.64)*0.01*(5/150)*1000*1000*(1/2))
PE_n <- neg_dat$`PE 15:0_18:1(d7)`
PE_n_pmol <- ((5.3/711.03)*0.01*(5/150)*1000*1000*(1/2))
PC_n <- neg_dat$`PC 15:0_18:1(d7)`
PC_n_pmol <- ((150.6/753.11)*0.01*(5/150)*1000*1000*(1/2))
PG_n <- neg_dat$`PG 15:0_18:1(d7)`
PG_n_pmol <- ((26.7/764.02)*0.01*(5/150)*1000*1000*(1/2))
PI_n <- neg_dat$`PI 15:0_18:1(d7)`
PI_n_pmol <- ((8.5/847.13)*0.01*(5/150)*1000*1000*(1/2))
PS_n <- neg_dat$`PS 15:0_18:1(d7)`
PS_n_pmol <- ((3.9/777.02)*0.01*(5/150)*1000*1000*(1/2))
SM_n <- neg_dat$`SM d18:1_18:1(d9)`
SM_n_pmol <- ((29.6/738.12)*0.01*(5/150)*1000*1000*(1/2))
```

```{r}
neg_dat[,grepl("^AcylGlcADG", names(neg_dat))] <- 
  (neg_dat[,grepl("^AcylGlcADG", names(neg_dat))]/PI_n)*PI_n_pmol
neg_dat[,grepl("^Cer" , names(neg_dat))] <- 
  (neg_dat[,grepl("^Cer" , names(neg_dat))]/SM_n)*SM_n_pmol
neg_dat[,grepl("^CL", names(neg_dat))] <- 
  (neg_dat[,grepl("^CL", names(neg_dat))]/PG_n)*PG_n_pmol
neg_dat[,grepl("^EtherOxPC", names(neg_dat))] <- 
  (neg_dat[,grepl("^EtherOxPC", names(neg_dat))]/PE_n)*PE_n_pmol
neg_dat[,grepl("^EtherPC", names(neg_dat))] <- 
  (neg_dat[,grepl("^EtherPC", names(neg_dat))]/PC_n)*PC_n_pmol
neg_dat[,grepl("^EtherPE", names(neg_dat))] <- 
  (neg_dat[,grepl("^EtherPE", names(neg_dat))]/PE_n)*PE_n_pmol
neg_dat[,grepl("^FA", names(neg_dat))] <- 
  (neg_dat[,grepl("^FA", names(neg_dat))]/LPE_n)*LPE_n_pmol
neg_dat[,grepl("^GlcADG", names(neg_dat))] <- 
  (neg_dat[,grepl("^GlcADG", names(neg_dat))]/PI_n)*PI_n_pmol
neg_dat[,grepl("^GM3" , names(neg_dat))] <- 
  (neg_dat[,grepl("^GM3" , names(neg_dat))]/SM_n)*SM_n_pmol
neg_dat[,grepl("^HBMP", names(neg_dat))] <- 
  (neg_dat[,grepl("^HBMP", names(neg_dat))]/PG_n)*PG_n_pmol
neg_dat[,grepl("^HexCer" , names(neg_dat))] <- 
  (neg_dat[,grepl("^HexCer" , names(neg_dat))]/SM_n)*SM_n_pmol
#neg_dat[,grepl("^LPA", names(neg_dat))] <- 
  #(neg_dat[,grepl("^LPA", names(neg_dat))]/LPE_n)*LPE_n_pmol
neg_dat[,grepl("^LPG", names(neg_dat))] <- 
  (neg_dat[,grepl("^LPG", names(neg_dat))]/PG_n)*PG_n_pmol
neg_dat[,grepl("^LPI", names(neg_dat))] <- 
  (neg_dat[,grepl("^LPI", names(neg_dat))]/PI_n)*PI_n_pmol
neg_dat[,grepl("^LPS", names(neg_dat))] <- 
  (neg_dat[,grepl("^LPS", names(neg_dat))]/PS_n)*PS_n_pmol
neg_dat[,grepl("^MGDG", names(neg_dat))] <- 
  (neg_dat[,grepl("^MGDG", names(neg_dat))]/PI_n)*PI_n_pmol
neg_dat[,grepl("^OxPC", names(neg_dat))] <- 
  (neg_dat[,grepl("^OxPC", names(neg_dat))]/PE_n)*PE_n_pmol
neg_dat[,grepl("^OxPE", names(neg_dat))] <- 
  (neg_dat[,grepl("^OxPE", names(neg_dat))]/PE_n)*PE_n_pmol
neg_dat[,grepl("^OxPI", names(neg_dat))] <- 
  (neg_dat[,grepl("^OxPI", names(neg_dat))]/PI_n)*PI_n_pmol
neg_dat[,grepl("^PEtOH", names(neg_dat))] <- 
  (neg_dat[,grepl("^PEtOH", names(neg_dat))]/PE_n)*PE_n_pmol
neg_dat[,grepl("^PMeOH", names(neg_dat))] <- 
  (neg_dat[,grepl("^PMeOH", names(neg_dat))]/PE_n)*PE_n_pmol
neg_dat[,grepl("^SHexCer" , names(neg_dat))] <- 
  (neg_dat[,grepl("^SHexCer" , names(neg_dat))]/SM_n)*SM_n_pmol
neg_dat[,grepl("^SQDG" , names(neg_dat))] <- 
  (neg_dat[,grepl("^SQDG" , names(neg_dat))]/PI_n)*PI_n_pmol
```

```{r}
neg_dat[,grepl("^LPC", names(neg_dat))] <- 
  (neg_dat[,grepl("^LPC", names(neg_dat))]/LPE_n)*LPE_n_pmol
neg_dat[,grepl("^LPE", names(neg_dat))] <- 
  (neg_dat[,grepl("^LPE", names(neg_dat))]/LPE_n)*LPE_n_pmol
#neg_dat[,grepl("^PA", names(neg_dat))] <- 
 # (neg_dat[,grepl("^PA", names(neg_dat))]/PA_n)
neg_dat[,grepl("^PC", names(neg_dat))] <- 
  (neg_dat[,grepl("^PC", names(neg_dat))]/PE_n)*PC_n_pmol
neg_dat[,grepl("^PE", names(neg_dat))] <- 
  (neg_dat[,grepl("^PE", names(neg_dat))]/PE_n)*PE_n_pmol
neg_dat[,grepl("^PG", names(neg_dat))] <- 
  (neg_dat[,grepl("^PG", names(neg_dat))]/PG_n)*PG_n_pmol
neg_dat[,grepl("^PI", names(neg_dat))] <- 
  (neg_dat[,grepl("^PI", names(neg_dat))]/PI_n)*PI_n_pmol
neg_dat[,grepl("^PS", names(neg_dat))] <- 
  (neg_dat[,grepl("^PS", names(neg_dat))]/PS_n)*PS_n_pmol
neg_dat[,grepl("^SM" , names(neg_dat))] <- 
  (neg_dat[,grepl("^SM" , names(neg_dat))]/SM_n)*SM_n_pmol
```

#Remove lipids that did not have a representative standard to normalize to
```{r}
neg_dat <- neg_dat[,!grepl(c("PA"),colnames(neg_dat))]
```

#Combine pos and neg lipids
Some lipids are identified in both positive and negative ionization. This is most common with phospholipids. 
First, these common lipids must be identified. Then, I drop the common lipid with a lower signal from either the positive or negative list. 

It makes the most sense to keep the one with higher signal because the peak integration was likely better.
```{r}
shared <- intersect(colnames(pos_dat)[5:ncol(pos_dat)], 
                    colnames(neg_dat)[5:ncol(neg_dat)])
n_s <- neg_dat[,shared]
p_s <- pos_dat[,shared]
negMode = as.numeric()
posMode = as.numeric()
for (i in 1:ncol(n_s)){
  negMode[i] = round(summary(n_s[[i]])[3],4)
  posMode[i] = round(summary(p_s[[i]])[3],4)
}
common <- cbind(colnames(n_s), negMode, posMode)
common <- as_tibble(common)
common[,2:3] <- sapply(common[,2:3], as.numeric)
NegHigh <- common$V1[common[[2]] > common[[3]]]  # higher intensity in negative mode
PosHigh <- common$V1[common[[3]] > common[[2]]]  # higher intensity in positive mode
# Remember to exclude the variables higher in negative mode from the POSITIVE data frame! And vice-versa
neg_final <- neg_dat[,!names(neg_dat) %in% PosHigh]
pos_final <- pos_dat[,!names(pos_dat) %in% NegHigh]
# Double check no common variables remain (IS is fine)
shared_test <- intersect(colnames(neg_final)[10:ncol(neg_final)], 
                    colnames(pos_final[10:ncol(pos_final)]))
print(shared_test)
# Final merged dataset
pos_merge <- pos_final[,c(1,6:ncol(pos_final))]
Lipids_Final <- merge(neg_final, pos_merge, by = "Name")
Lipids_Final <- Lipids_Final %>% 
                  tibble %>% 
                  select(colnames(Lipids_Final)[1:5], sort(colnames(.)))
```

# Normalize to starting weight
```{r}
IS_rm <- colnames(select(Lipids_Final, matches("(d7)")))
IS_rm  <- append(IS_rm, colnames(select(Lipids_Final, matches("(d9)"))))
Lipids_Final <- Lipids_Final[,!names(Lipids_Final) %in% IS_rm]
Lipids_Final_norm <- Lipids_Final
Lipids_Final_norm[,5:ncol(Lipids_Final_norm)] <- Lipids_Final[,5:ncol(Lipids_Final)]/Lipids_Final$weight
Lipids_Final_norm[is.na(Lipids_Final_norm)] <- 0
Lipids_Final_norm$Name <- as.numeric(Lipids_Final_norm$Name)
Lipids_Final_norm  <- Lipids_Final_norm[order(Lipids_Final_norm$Name),]
#colnames(Lipids_Final_norm) <- gsub("'\\.'.*", "", colnames(Lipids_Final_norm)) #remove the duplicate annotations
write_csv(Lipids_Final_norm, "cleanup/ether_height_pmol_mg_tissue.csv")
rm(list=ls())
```

# For metaboanalyst 
```{r}
library(dplyr)
MA <- Lipids_Final_norm
#MA <- MA %>% select(-contains("RT_"))
MA <- MA[,-4] #remove weight column
MA_2 <- MA %>% unite("Condition",2:3,remove=T)
write_csv(MA_2, "cleanup/combined_pmol_mg_tissue_MA.csv")
rm(list=ls())
```

